Search for Spatially Extended Fermi-LAT Sources Using Two Years of Flight Data

### 2011-08-15 extended_catalog/paper/extended_catalog.tex
working on interpretation

\usepackage{rotating}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage{url}

@@

% some units
\newcommand{\kev}{\text{Kev}\xspace}
\newcommand{\mev}{\text{MeV}\xspace}
\newcommand{\gev}{\text{GeV}\xspace}
\newcommand{\tev}{\text{TeV}\xspace}
\newcommand{\sr}{\text{sr}\xspace}
\newcommand{\s}{\text{s}\xspace}
\newcommand{\ph}{\text{ph}\xspace}
\newcommand{\cm}{\text{cm}\xspace}
\renewcommand{\sec}{\text{s}\xspace}
\newcommand{\tsext}{{\ensuremath{\text{TS}_{\text{ext}}}}\xspace}
\newcommand{\tsinc}{\ensuremath{\text{TS}_{\text{inc}}}\xspace}
\newcommand{\loglikelihood}{\ensuremath{LL}\xspace}
\newcommand{\chandra}{\text{{\em Chandra}}\xspace}

@@

\todo[inline]{Remove from all figures - "more infomration can be foudn in"}

@@

@@

todo[inline]{This text needs to be fixed}

Figure~\ref{diff_factor_sensitivity} shows how the LAT's detection
threshold increases with increasing background. We compute the threshold
using 1, 10, and 100 times the Sreekumar-like isotropic background. The
latter factor approximately represents the intensity of the Galactic
diffuse emission near the Galactic center.  The figure shows the threshold
using 1-100 \gev photons and also
using only 10-100 \gev photons. We
are less sensitive to extended sources with higher background. Also,
given the same flux we are more sensitive to the source if those
photons have a higher energies.

Overlaid on figure~\ref{diff_factor_sensitivity} are the LAT
detected extended sources from table~\ref{known_extended_sources} and
table~\ref{new_ext_srcs_table}.  The extended sources are color coded
based upon whether the galactic diffuse and isotropic emission model is
closer to 1, 10, or 100 times the Sreekumar-like isotropic spectrum.

Figure 7 shows the flux threshold as a function of source extension for
different background levels (1x, 10x, and 100x the
nominal background), different spectral indices, and two
different energy bands (1-100 \gev and 10-100 \gev).
For a fixed 1-100 \gev or 10-100 \gev
flux, the LAT's detection threshold is only weakly depends upon the
spectral index.  This effect is most pronounced when using only photons
with energies above 10 \gev. The extension thresholds are tabulated in
table~\ref{all_sensitivity_table}.

@@

We first present on our analysis of the twelve extended sources
included in 2FGL.
(\cite{second_cat}).
The six extended SNRs IC443, W28, W30, W44, W51C, and the
Cygnus Loop were included in 2FGL.  IC443, W28, W44, W51C were
first reported in (\cite{ic443,w28,w44,w51c}).  Detailed papers by the LAT
collaboration studying W30 and the Cygnus Loop are still in preparation.
\gev emission from W30 was also studied in (\cite{castro_and_slane_2010}).
Using 1-100 \gev photons, our analysis significantly detected the
extension of all six SNRs.
% 2 galaxies

@@

Following 1FGL, the LAT collaboration published a list called 1LAC
of sources that
had a high probability association with AGN (\cite{first_agn_cat}).
1FGL had 709 1FGL sources associated with 671 distinct AGN at high
latitude ($|b|>10\deg$).  An upcoming publication by the LAT collaboration
will produce a second list 2LAC. 2LAC 
uses the 2FGL catalog and associates 1017 2FGL sources
with AGN.  To avoid systematic problems with AGN classification, we
selected 815 out of the 1017 AGN which made it into the clean AGN sub-sample.  An AGN
association is considered clean only if it has a high probability
of association $P\ge 80\%$, if it is the only AGN associated with the
catalog source, and if there are no flags on the source in 2FGL. These
last two conditions are important for our analysis. Source
confusion may look like a spatially extended source and flagged catalog
sources may correlate with unmodeled structure in the diffuse emission.

Of the 815 clean AGN, we select 775 of these 2FGL sources which
are significant above 1 \gev and fit each of them for extension.
A histogram of the \tsext values computed for these AGN is
shown in figure~\ref{agn_ts_ext}. Overlaid on the plot is the
$\chi^2/2$ distribution of equation~\ref{ts_ext_distribution}.
The \tsext distribution for AGN shows good agreement with the
theoretical distribution.  Two sources had $\tsext>9$.  One was due
to incorrectly automatically removing a nearby catalog sources (see
section~\ref{extended_source_search_method}) and the other was due to a
failure of convergence of the point hypothesis.  This result demonstrates
that we can use \tsext as a measure of the statistical significance of
the detection of the extension of a source.

@@

The results of our investigation of systematic uncertainties of this
measurement are presented in table~\ref{dual_localization_results}
and~\ref{alt_diff_model_results}.  Table~\ref{dual_localization_results}
shows a comparison between the fit with a single extended source
hypothesis and the fit assuming the emission originates from two
independent sources.  Table~\ref{alt_diff_model_results} shows the
results of the extension fit using variations of the PSF and the galactic
diffuse model described in section~\ref{systematic_errors_on_extension}.
There is good agreement between \tsext and the fit size using
the standard analysis, the alternative diffuse models, and the alternative PSF.
This suggests that the sources are robust against features in the diffuse
model and uncertainties in the angular resolution.

@@

This extended source is spatially coincident with the SNR Puppis A.
Overlaid on figure~\ref{1FGL_J0823.3-4248} are ROSAT contours of Puppis
A (\cite{rosat_puppis_a}). There is good agreement between of the X-ray
shape with the fit size in \gev.

\todo[inline]{Add interpretations of Puppis A.}
\todo[inline]{Markus says to add an SED.}

@@

This source is in a region of remarkably complicated diffuse emission.
Even though it is 16\deg from the Galactic plane, the source is on top
of the core of the Ophiuchus molecular cloud which contains massive
star-forming regions that are bright in the infrared.  We have overlaid
on figure~\ref{1FGL_J1628.6-2419c} contours of the 100 micrometer infared
image of this region measured by IRAS. There is good spatial overlap
with the \gev emission (\cite{iras_rho_ophiuci}).

The region also has abundant molecular and atomic gas traced by CO
and HI but also plenty of dark gas, the kind that we find only by its
association with dust emission (\cite{isabelle_dark_gass}). Embedded
star-forming regions make it even more challenging to measure the column
density of dust.  Therefore this source might represent an inadequacy
in the diffuse $\gamma$-ray model

@@

HESS\,J1616-508 is located in the region of two SNRs, RCW103
(G332.4-04) and Kes 32 (G332.4+0.1) but is not spatially coincident
with either of them (\cite{hess_plane_survey}).  HESS\,J1616-508 is near
three pulsars PSR J1614-5048, PSR J1616-5109, and PSR J1617-5055
but only PSR J1617-5055 is energetically favored as being the \tev
emitter (\cite{discovery_of_PSR_J1617-5055,integral_HESS_J1616-508}).
\cite{hess_plane_survey} speculated that a PWN powered by this young
pulsar could be responsible for the emission of HESS\,J1616-508.
Because HESS\,J1616-508 is 9' away from PSR J1617-5055, this would
require an assymetric X-ray PWNe to power the \tev emission. However,
\chandra ACIS observations revealed only an underluminous PWN of
about $\approx1'$ size around the pulsar which was not oriented
towards the \tev emission, rendering this association as uncertain
\cite{discovery_of_pwn_for_PSR_J1617-5055}.  No other promising
counterparts were observed in observations of x-ray and soft gamma-ray
emission by Suzaku (\cite{suzakzu_HESS_J1616-508}), Swift/XRT,
IBIS/ISGRBI, BeppoSAX and XMM-Newton (\cite{integral_HESS_J1616-508}).
Finally, \cite{discovery_of_pwn_for_PSR_J1617-5055} found diffuse
emission towards the center of HESS\,J1616-508 using archival radio and
infared observations.  Deeper observations will likely be necessary to
understand this region.

@@

\todo[inline]{Swift observations}

@@

Swift observations the region and resolved six X-ray sources.\todo{citation}

\todo[inline]{Emission due to SNR}

\todo[inline]{Emission could be due to a PWN}

\todo[inline]{Emission due to Massive Star}

\todo[inline]{Add interpretations of HESS\,J1614-518. Stefan says to
Suzaku observations.}

@@

\cite{hess_plane_survey} argued that HESS\,J1632-478 is
positionally coincident with the hard X-ray source IGR\,J1632-4751
observed by INTEGRAL (\cite{Igr_J16320-4751_circ}), XMM-Newton
(\cite{xmm_newton_IGR_J16320-4751}), and ASCI (\cite{asca_plane_survey}).
But it is suspected to be an galactic X-Ray Binary so the \gev and \tev
extension disfavors the association.

Further observations by XMM-Newton reveal extended emission of size
$\approx32''x\approx15''$ and inside point-like emission coincident with
the peak H.E.S.S. emimssion (\cite{hess_j1632_478_xmm_newton}).  They
found in archival MGPS-2 data a spatially coincident extended radio source
\cite{most_survey_galactic_plane}.  The positional match argues for a
single synchrotron processes producing X-ray and \gev and \tev radiation,
likely due to a PWNe.  The increased \gev and \tev size compared
to the X-ray size has previously been observed in several aging PWNe
including HESS\,J1825-137 (\cite{hess_j1825_xmm_newton,hess_j1825_hess},
HESS\,J1640-465 (\cite{hess_plane_survey,xmm_newton_hess_j_1640-466}),
and Vela X (\cite{vela_x_rosat,vela_x_hess}).  This can be explained by
a different synchrotron cooling time for the electrons producing X-rays
and \tev gamma rays.

@@

This source is spatially coincident with the Gamma Cygni SNR.  Contours of
Gamma Cygni at 408MHz from the Canadian Galactic Plane Survey are overlaid
on figure~\ref{1FGL_J2020.0+4049} and there is good spatial agreement
with the LAT emission (\cite{canadian_galactic_plane_survey}).

\todo[inline]{
Add interpretations paragraph.  Describe how this source will be more
carefully analyzed in ``A cocoon of freshly accelerated cosmic rays
detected by Fermi in the Cygnus superbubble''}

@@

Figure~\ref{gev_vs_tev_histogram} compares the size of the 21 extended
LAT sources to the 42 extended H.E.S.S. sources.  Because of the large
field of view and all sky coverage, the LAT can more easily measure
larger sources including the LMC and SMC.  On the other hand, the smaller
angular resolution of Air Cherenkov detectors allows them to measure a
population of extended sources below our sensitivity limit (currently at
about $\approx0.2\deg$).  Fermi has a five year nominal livetime with a
goal of ten years of operation.  As figure~\ref{time_sensitivity} showes,
our sensitivity to these smaller sources will improve by a factor greater
than $\sqrt{t}$.  With increased exposure, the LAT will likely begin to
resolve these smaller \tev sources.

@@

\begin{figure}
  \begin{center}
    \includegraphics{mc_plots/compare_disk_gauss.pdf}
    % this plot came from /u/gl/lande/work/fermi/extended_catalog/2FGL/plots_for_paper/compare_disk_gauss/v2/compare_disk_gauss.pdf
    \end{center}
    \caption{
    A comparison of a Gaussian and disk like surface brightness profile
    of extended sources before and after convolving with the PSF for two
    energy ranges.  The black line is the PSF that would be observed
    for a power-law source of spectral index 2. The dashed red line
    is the brightness profile of a Gaussian with $\rsixeight=0.5\deg$.
    The dashed blue line is the brightness profile for a uniform disk
    with $\rsixeight=0.5\deg$.  The solid red and blue lines show the
    convolution of the Gaussian and uniform disk brightness profiles
    with the LAT PSF.
    }\label{compare_disk_gauss}
  \end{figure}

\begin{deluxetable}{rrrr}
\tabletypesize{\normalsize}
\tablecaption{Monte Carlo Validation Parameters
\label{ts_ext_num_sims}
}
\tablewidth{0pt}
\tablehead{

@@

\tablecomments{
    the spectral models used in the Monte Carlo validation of \pointlike.
    For each model, the number of statistically independent simulations
    and the average value of \ts is tabulated.  The models spans
    representative spectral parameters.
}

@@

% \begin{figure}
%   \begin{center}
%     \includegraphics{mc_plots/diff_factor_sensitivity.pdf}
%     % this plot came from /u/gl/lande/work/fermi/extended_catalog/monte_carlo/sensitivity/v12/plot/plot_vs_diff_factor.py
%     \end{center}
%     \caption{
%     This shows the detection threshold to a uniform disk extended
%     source for varying background levels
%     and energy ranges.  The plots have the same meaning as in
%     figure~\ref{index_sensitivity} and the simulations occur over the
%     same time range.  The solid blue line is the detection threshold against a
%     Sreekumar-like isotropic background. The green dashed curve is the
%     threshold against 10 times the isotropic background and the dash
%     dotted red curve is the threshold against 100 times the isotropic
%     background.  The latter factor approximately represents the intensity
%     of the Galactic diffuse emission near the Galactic center.  Plot (a)
%     is the threshold using 1-100\gev photons while plot (b) is the threshold using
%     10-100\gev photons. Unlike figure~\ref{index_sensitivity},
%     the detection threshold was calculated only for a power-law source of spectral
%     index 2. The flux is quoted only in the selected energy range which
%     weakens the dependence of the detection threshold on spectral index.
%     These plots show that we are less sensitive to extended sources in
%     regions with higher background and (given the same number of photons)
%     more sensitive to harder sources.  Overlaid on this plot are the
%     extended sources found in this extended source search. The extended
%     sources are color coded based upon whether the integral galactic
%     diffuse and isotropic diffuse emission model is closer to 1x, 10x,
%     or 100x the Sreekumar-like isotropic spectrum.  The red point
%     in plot (a) below the red line is RX J1713.7-3946 which is harder
%     ($\gamma\approx1.5$) and has diffuse emission only 53 times larger
%     than the isotropic emission.  The green point in plot (b) below
%     the green line is MSH 15-52 which was found by our analysis to
%     have $\tsext=6.5$.  More information about this plot is presented
%     in section~\ref{extension_sensitivity}.
%     }\label{diff_factor_sensitivity}
%   \end{figure}

\begin{figure}
  \begin{center}
    \includegraphics{mc_plots/all_sensitivity.pdf}
    % this plot came from /u/gl/lande/work/fermi/extended_catalog/monte_carlo/sensitivity/v12/plot/plot_all.py
    \end{center}
    \caption{the LAT detection threshold for four spectral indices and
    three background (1x, 10x, and 100x the Sreekumar-like isotropic
    spectrum) for a two-year exposure. The left plots are the detection
    threshold when using 1-100 \gev photons and the right plots are
    the detection threshold when using 10-100 \gev photons.  The flux
    is integrated only in the selected energy range.  Overlaid on this
    plot are the LAT detected extended sources placed by the nearby
    Galactic diffuse emission and color coded by their spectral idnex.
    The black point in plot (d) below the green line is MSH 15-52 which
    was found by our analysis to have $\tsext=6.5$.  
    }\label{all_sensitivity} 
  \end{figure}

@@

  \begin{sidewaystable}\scriptsize
    \begin{centering}
      \begin{tabular}{r|r|rrrrrrrrrrrrrrrrrrrr}

@@

        \multicolumn{22}{c}{$ 1 \gev < E < 100 \gev$} \\
        $\gamma$ &       BG &  $0.1\deg$ &  $0.2\deg$ &  $0.3\deg$ &  $0.4\deg$ &  $0.5\deg$ &  $0.6\deg$ &  $0.7\deg$ &  $0.8\deg$ &  $0.9\deg$ &    $1\deg$ &  $1.1\deg$ &  $1.2\deg$ &  $1.3\deg$ &  $1.4\deg$ &  $1.5\deg$ &  $1.6\deg$ &  $1.7\deg$ &  $1.8\deg$ &  $1.9\deg$ &    $2\deg$ \\
             1.5 &       x1 &      148.3 &       23.2 &       11.2 &        8.0 &        7.1 &        6.9 &        6.8 &        6.8 &        7.0 &        7.3 &        7.5 &        7.7 &        8.0 &        8.3 &        9.1 &        9.8 &        9.0 &        9.1 &        8.8 &       10.2 \\
                 &      x10 &      146.9 &       29.0 &       18.5 &       15.1 &       15.4 &       15.0 &       14.7 &       15.6 &       16.6 &       17.3 &       18.0 &       19.1 &       20.7 &       22.5 &       23.6 &       24.3 &       21.2 &       22.8 &       23.3 &       23.6 \\
                 &     x100 &      184.3 &       55.1 &       43.0 &       41.0 &       40.7 &       41.7 &       40.3 &       40.8 &       43.0 &       43.6 &       38.3 &       39.5 &       40.1 &       38.5 &       36.5 &       35.8 &       36.5 &       38.1 &       36.6 &       37.1 \\
               2 &       x1 &      327.5 &       43.1 &       18.7 &       13.3 &       11.2 &       10.4 &       10.1 &       10.1 &       10.2 &       10.4 &       10.6 &       10.9 &       11.1 &       11.5 &       12.3 &       12.6 &       13.0 &       13.4 &       13.8 &       14.2 \\
                 &      x10 &      336.1 &       55.3 &       32.2 &       27.4 &       26.3 &       25.3 &       25.5 &       25.8 &       27.2 &       26.6 &       27.6 &       28.5 &       30.1 &       29.9 &       30.8 &       31.8 &       31.9 &       34.2 &       34.2 &       36.1 \\
                 &     x100 &      409.6 &      128.1 &       88.6 &       76.2 &       72.7 &       71.4 &       67.8 &       65.2 &       63.6 &       63.7 &       62.0 &       63.1 &       62.4 &       61.1 &       56.0 &       58.4 &       60.1 &       62.1 &       59.5 &       55.9 \\
             2.5 &       x1 &      622.9 &       75.0 &       29.6 &       19.0 &       15.5 &       13.6 &       12.8 &       12.7 &       12.4 &       12.5 &       12.6 &       12.9 &       12.8 &       13.1 &       13.4 &       13.7 &       14.1 &       14.6 &       15.0 &       15.7 \\

>>>>

%\usepackage{rotating}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage{url}

@@

% some units
\newcommand{\kev}{\text{Kev}\xspace}
\newcommand{\mev}{\text{MeV}\xspace}
\newcommand{\gev}{\text{GeV}\xspace}
\newcommand{\tev}{\text{TeV}\xspace}
\newcommand{\sr}{\text{sr}\xspace}
\newcommand{\s}{\text{s}\xspace}
\newcommand{\ph}{\text{ph}\xspace}
\newcommand{\cm}{\text{cm}\xspace}
\renewcommand{\sec}{\text{s}\xspace}
\newcommand{\tsext}{{\ensuremath{\text{TS}_{\text{ext}}}}\xspace}
\newcommand{\tsinc}{\ensuremath{\text{TS}_{\text{inc}}}\xspace}
\newcommand{\loglikelihood}{\ensuremath{LL}\xspace}
\newcommand{\chandra}{\text{{\em Chandra}}\xspace}
\newcommand{\rosat}{\text{{\em ROSAT}}\xspace}
\newcommand{\asca}{\text{{\em ASCA}}\xspace}
\newcommand{\einstein}{\text{{\em Einstein}}\xspace}
\newcommand{\xmmnewton}{\text{{\em XMM-Newton}}\xspace}

@@

@@

\todo[inline]{Ask Stefan/Markus if they think we should get rid of
this note about what point soruces in the plane.}

@@

Figure~\ref{all_sensitivity}
shows the flux threshold as a function of source extension for
different background levels (1x, 10x, and 100x the
nominal background), different spectral indices, and two
different energy bands (1-100 \gev and 10-100 \gev).
The detection threshold is higher for sources in regions
of higher background.
For a fixed 1-100 \gev or 10-100 \gev
flux, the LAT's detection threshold is only weakly depends upon the
spectral index.  
Overlaid on figure~\ref{all_sensitivity} are the LAT
detected extended sources from table~\ref{known_extended_sources} and
table~\ref{new_ext_srcs_table}.  
This effect is most pronounced when using only photons
with energies above 10 \gev. The extension thresholds are tabulated in
table~\ref{all_sensitivity_table}.

Finally, figure~\ref{time_sensitivity} shows the LAT's projected detection
threshold to extension after 10 years against 10 times the isotropic
background. This background is representative of the background near
the Galactic plane.  For small extended sources, our detection threshold
improves better than $\sqrt{\text{t}}$ because at high energies, where we
are most sensitive to extension, the background levels are in the Poisson
instead of Gaussian regime.  For large extended sources, the relevant
background is larger and the improvement is closer to $\sqrt{\text{t}}$.

@@

We first present on our analysis of the twelve extended sources
included in 2FGL.
(\cite{second_cat}).
The six extended SNRs IC443, W28, W30, W44, W51C, and the
Cygnus Loop were included in 2FGL.  IC443, W28, W44, W51C, and the
Cygnus Loop were
first reported in (\cite{ic443,w28,w44,w51c,cygnus_loop_lat}).  A
detailed publication by the LAT
collaboration about W30 is still in preparation.
\gev emission from W30 was also studied in (\cite{castro_and_slane_2010}).
Using 1-100 \gev photons, our analysis significantly detected the
extension of all six SNRs.
% 2 galaxies

@@

Following 1FGL, the LAT collaboration published a list called 1LAC
of sources that
had a high probability association with AGN (\cite{first_agn_cat}).
1FGL had 709 1FGL sources associated with 671 distinct AGN at high
latitude ($|b|>10\deg$).  
Using two years of data and 2FGL, the LAT collaboraiton
associated 1016 2FGL sources
with AGN (\cite{second_agn_cat}).  
To avoid systematic problems with AGN classification, we
selected 885 out of the 1016 AGN which made it into the clean AGN sub-sample.  An AGN
association is considered clean only if it has a high probability
of association $P\ge 80\%$, if it is the only AGN associated with the
catalog source, and if there are no flags on the source in 2FGL. These
last two conditions are important for our analysis. Source
confusion may look like a spatially extended source and flagged catalog
sources may correlate with unmodeled structure in the diffuse emission.

Of the 885 clean AGN, we select 783 of these 2FGL sources which
are significant above 1 \gev and fit each of them for extension.
A histogram of the \tsext values computed for these AGN is
shown in figure~\ref{agn_ts_ext}. Overlaid on the plot is the
$\chi^2/2$ distribution of equation~\ref{ts_ext_distribution}.
The \tsext distribution for AGN shows good agreement with the
theoretical distribution.  Two sources had $\tsext>10$.  One was due
to incorrectly automatically removing a nearby catalog sources (see
section~\ref{extended_source_search_method}) and the other was due to a
failure of convergence of the point hypothesis.  This result demonstrates
that we can use \tsext as a measure of the statistical significance of
the detection of the extension of a source.

@@

The results of our investigation of systematic uncertainties of this
measurement are presented in 
table~\ref{alt_diff_model_results}.  It
shows a comparison between the fit with a single extended source
hypothesis and the fit assuming the emission originates from two
independent sources and  the
results of the extension fit using variations of the PSF and the galactic
diffuse model described in section~\ref{systematic_errors_on_extension}.
There is good agreement between \tsext and the fit size using
the standard analysis, the alternative diffuse models, and the alternative PSF.
This suggests that the sources are robust against features in the diffuse
model and uncertainties in the angular resolution.

@@

This extended source is spatially coincident with the middle-aged
SNR Puppis A.  Puppis A has been studied in detail in radio
(\cite{puppis_a_vla}, and references therein).  X-ray observations
have been performed by a variety of instruments 
(\cite{einstein_puppis_a,rosat_puppis_a,chandra_puppis_a,xmm_newton_puppis_a,suzaku_puppis_a})
The highest resolution image of the entire source was obtained
by \rosat and contours corresponding to this
image are overlaid on figure~\ref{1FGL_J0823.3-4248} and match the
inferred \gev size (\cite{rosat_puppis_a}).

Puppis A's age was estimated to be $\approx3700$ years (\cite{winkler_1988}) and
its distance was estimated at 2.2 kpc (\cite{ reynoso_1995,reynoso_2003}).
Assuming this distance, the \gev luminosity of Puppis A between 1 and
100 \gev is $\approx 3\times 10^{34}$ ergs/\s.

Molecular clouds have not been detected ajacent to Puppis A (\cite{puppis_a_co})
Similarly, the luminsoity is lower than that of other LAT
detected SNRs interacting with molecular clouds (\cite{\dots}),.
The LAT emission could therefore be explained by the interaciton of acclerated
particles by interstellar matter or fields in the regions, similar
to what was suggested for the Cygnus Loop SNR (\cite{cygnus_loop_lat}).

Puppis A has not been observed to direct
% 1 - 100 GeV
%      Norm      
%      Index     
% Flux = PowerLaw(norm=6.21e-14,index=2.22,e0=10000).i_flux(cgs=True,emin=1e3,emax=1e5,e_weight=1)
%      = 4.7804242225163421e-11 ergs/cm^2/s
% Luminosity =  Flux * 4*pi*r^2 = 4.78e-11 ergs/cm^2/s * 4*pi * (2.2 kpc )**2 in ergs/s
%            = 2.8e34 ergs/s from 1-100 gev

\todo[inline]{Make point about how this source
is like Cygnus Loop. Discuss with Yasunobu.  Main points about CO
Emission / moelcular cloud interactions (\cite{co_eastern_puppis_a}).
Can we say anything about emission could be hadronic or leptonic models?
Markus says to add SEDs. Should we?
}

@@

This source is in a region of remarkably complicated diffuse emission.
Even though it is 16\deg from the Galactic plane, the source is on
top of the core of the Ophiuchus molecular cloud which contains massive
star-forming regions that are bright in the infrared.  The region also has
abundant molecular and atomic gas traced by CO and HI but also plenty of
dark gas, the kind that we find only by its association with dust emission
(\cite{isabelle_dark_gass}). Embedded star-forming regions make it even
more challenging to measure the column density of dust.  Infared and CO
contours are overlaid on figure~\ref{1FGL_J1628.6-2419c} and there is good
spatial correlation with the \gev emission (\cite{iras_rho_ophiuci}).
Therefore this source might represent an inadequacy in the diffuse
$\gamma$-ray model

@@

HESS\,J1616-508 is located in the region of two SNRs, RCW103
(G332.4-04) and Kes 32 (G332.4+0.1) but is not spatially coincident
with either of them (\cite{hess_plane_survey}).  HESS\,J1616-508 is near
three pulsars PSR J1614-5048, PSR J1616-5109, and PSR J1617-5055
but only PSR J1617-5055 is energetically favored as being the \tev
emitter (\cite{discovery_of_PSR_J1617-5055,integral_HESS_J1616-508}).
\cite{hess_plane_survey} speculated that a PWN powered by this young
pulsar could be responsible for the emission of HESS\,J1616-508.
Because HESS\,J1616-508 is 9' away from PSR J1617-5055, this would
require an assymetric X-ray PWNe to power the \tev emission. However,
\chandra ACIS observations revealed only an underluminous PWN of
about $\approx1'$ size around the pulsar which was not oriented
towards the \tev emission, rendering this association as uncertain
\cite{discovery_of_pwn_for_PSR_J1617-5055}.  No other promising
counterparts were observed in observations of x-ray and soft gamma-ray
emission by Suzaku (\cite{suzakzu_HESS_J1616-508}), Swift/XRT,
IBIS/ISGRBI, BeppoSAX and \xmmnewton (\cite{integral_HESS_J1616-508}).
Finally, \cite{discovery_of_pwn_for_PSR_J1617-5055} found diffuse
emission towards the center of HESS\,J1616-508 using archival radio and
infared observations.  Deeper observations will likely be necessary to
understand this region.

@@

@@

Swift observations the region and resolved six X-ray sources.

\todo[inline]{Emission due to SNR, PWN, Massive Star. Stefan
says to add Suzaku observations. Add note about SWIFT observations + citation.}

@@

\cite{hess_plane_survey} argued that HESS\,J1632-478 is
positionally coincident with the hard X-ray source IGR\,J1632-4751
observed by INTEGRAL (\cite{Igr_J16320-4751_circ}), \xmmnewton
(\cite{xmm_newton_IGR_J16320-4751}), and \asca (\cite{asca_plane_survey}).
But it is suspected to be an galactic X-Ray Binary so the \gev and \tev
extension disfavors the association.

Further observations by \xmmnewton reveal extended emission of size
$\approx32''x\approx15''$ and inside point-like emission coincident with
the peak H.E.S.S. emimssion (\cite{hess_j1632_478_xmm_newton}).  They
found in archival MGPS-2 data a spatially coincident extended radio source
\cite{most_survey_galactic_plane}.  The positional match argues for a
single synchrotron processes producing X-ray and \gev and \tev radiation,
likely due to a PWNe.  The increased \gev and \tev size compared
to the X-ray size has previously been observed in several aging PWNe
including HESS\,J1825-137 (\cite{hess_j1825_xmm_newton,hess_j1825_hess},
HESS\,J1640-465 (\cite{hess_plane_survey,xmm_newton_hess_j_1640-466}),
and Vela X (\cite{vela_x_rosat,vela_x_hess}).  This can be explained by
a different synchrotron cooling time for the electrons producing X-rays
and \tev gamma rays.

@@

This source is spatially coincident with the $\gamma$ Cygni SNR.  Contours of
$\gamma$ Cygni at 408MHz from the Canadian Galactic Plane Survey are overlaid
on figure~\ref{1FGL_J2020.0+4049} and there is good spatial agreement
with the LAT emission (\cite{canadian_galactic_plane_survey}).

\todo[inline]{Add some more information about $\gamma$ cygni.}

@@

Figure~\ref{gev_vs_tev_histogram} compares the size of the 21 extended
LAT sources to the 42 extended H.E.S.S. sources.  Because of the large
field of view and all sky coverage, the LAT can more easily measure
larger sources including the LMC and SMC.  On the other hand, the smaller
angular resolution of Air Cherenkov detectors allows them to measure a
population of extended sources below our sensitivity limit (currently at
about $\approx0.2\deg$).  Fermi has a five year nominal livetime with a
goal of ten years of operation.  As figure~\ref{time_sensitivity} shows,
our sensitivity to these smaller sources will improve by a factor greater
than $\sqrt{t}$.  With increased exposure, the LAT will likely begin to
resolve these smaller \tev sources.

@@

\clearpage
\begin{figure}
  \begin{center}
    \includegraphics{mc_plots/compare_disk_gauss.pdf}
    % this plot came from /u/gl/lande/work/fermi/extended_catalog/2FGL/plots_for_paper/compare_disk_gauss/v2/compare_disk_gauss.pdf
    \end{center}
    \caption{
    A comparison of a Gaussian and disk like surface brightness profile
    of extended sources before and after convolving with the PSF for two
    energy ranges.  The black line is the PSF that would be observed
    for a power-law source of spectral index 2. The dashed red line
    is the brightness profile of a Gaussian with $\rsixeight=0.5\deg$.
    The dashed blue line is the brightness profile for a uniform disk
    with $\rsixeight=0.5\deg$.  The solid red and blue lines show the
    convolution of the Gaussian and uniform disk brightness profiles
    with the LAT PSF.
    }\label{compare_disk_gauss}
  \end{figure}

\clearpage
\begin{deluxetable}{rrrr}
\tabletypesize{\normalsize}
\tablecaption{Monte Carlo Validation Parameters
\label{ts_ext_num_sims}
}
\tablewidth{0pt}
\tablehead{

@@

\tablecomments{
    the spectral models used in the Monte Carlo validation of \pointlike.
    For each model, the number of statistically independent simulations
    and the average value of \ts is tabulated.  The models spans
    representative spectral parameters.
}
\end{deluxetable}

@@

\clearpage
\begin{figure}
  \begin{center}
    \includegraphics{mc_plots/all_sensitivity.pdf}
    % this plot came from /u/gl/lande/work/fermi/extended_catalog/monte_carlo/sensitivity/v12/plot/plot_all.py
    \end{center}
    \caption{the LAT detection threshold for four spectral indices and
    three background (1x, 10x, and 100x the Sreekumar-like isotropic
    spectrum) for a two-year exposure. The left plots are the detection
    threshold when using 1-100 \gev photons and the right plots are
    the detection threshold when using 10-100 \gev photons.  The flux
    is integrated only in the selected energy range.  Overlaid on this
    plot are the LAT detected extended sources placed by their nearby
    Galactic diffuse emission and color coded by their spectral index.
    The black point in plot (d) below the green line is MSH 15-52 which
    was found by our analysis to have $\tsext=6.5$.  
    }\label{all_sensitivity} 
  \end{figure}

@@

\begin{deluxetable}{rrrrrrrrrrrrrrrrrrrrrrr}
\tabletypesize{\scriptsize}
\rotate
\tablewidth{0pt}
\tablecaption{Extension Detection Threshold
\label{all_sensitivity_table}
\tablehead{
\colhead{$\gamma$}&       
\colhead{BG}&
\colhead{$0.1$}&
\colhead{$0.2$}&
\colhead{$0.3$}&
\colhead{$0.4$}&
\colhead{$0.5$}&
\colhead{$0.6$}&
\colhead{$0.7$}&
\colhead{$0.8$}&
\colhead{$0.9$}&
\colhead{$1.0$}&
\colhead{$1.1$}&
\colhead{$1.2$}&
\colhead{$1.3$}&
\colhead{$1.4$}&
\colhead{$1.5$}&
\colhead{$1.6$}&
\colhead{$1.7$}&
\colhead{$1.8$}&
\colhead{$1.9$}&
\colhead{$2.0$}
\startdata
\multicolumn{22}{c}{E$>$1 \gev} \\

@@

     1.5 &       x1 &      148.1 &       23.3 &       11.3 &        8.0 &        7.2 &        6.9 &        6.7 &        6.8 &        7.1 &        7.4 &        7.6 &        7.9 &        8.1 &        8.5 &        9.2 &        9.9 &        9.1 &        9.2 &        9.0 &       10.3 \\
         &      x10 &      148.4 &       29.0 &       18.7 &       15.2 &       15.4 &       15.0 &       16.1 &       16.0 &       16.8 &       17.7 &       18.2 &       19.3 &       20.9 &       22.5 &       23.8 &       24.8 &       21.3 &       22.8 &       23.4 &       23.7 \\
         &     x100 &      186.8 &       55.0 &       43.4 &       40.7 &       41.0 &       41.8 &       40.9 &       40.9 &       42.7 &       43.6 &       38.4 &       39.9 &       40.6 &       38.4 &       36.9 &       36.3 &       37.1 &       38.8 &       37.2 &       37.6 \\
       2 &       x1 &      328.4 &       43.4 &       18.9 &       13.4 &       11.2 &       10.4 &       10.2 &       10.2 &       10.2 &       10.4 &       10.7 &       10.9 &       11.2 &       11.5 &       12.4 &       12.6 &       13.0 &       13.4 &       14.0 &       14.4 \\
         &      x10 &      341.0 &       55.9 &       32.3 &       27.6 &       26.5 &       25.4 &       25.6 &       25.9 &       27.4 &       26.8 &       27.8 &       28.7 &       29.8 &       30.1 &       31.0 &       31.5 &       31.7 &       34.0 &       34.3 &       35.9 \\
         &     x100 &      420.5 &      128.3 &       90.2 &       77.3 &       73.3 &       70.8 &       67.5 &       64.3 &       64.2 &       64.1 &       62.8 &       63.6 &       61.7 &       61.9 &       58.4 &       59.0 &       61.4 &       63.3 &       60.1 &       58.1 \\
     2.5 &       x1 &      627.1 &       75.6 &       29.8 &       19.3 &       15.5 &       13.5 &       12.8 &       12.6 &       12.5 &       12.5 &       12.6 &       12.9 &       12.9 &       13.1 &       13.5 &       13.7 &       14.3 &       14.8 &       15.2 &       15.8 \\
         &      x10 &      638.9 &       99.1 &       52.1 &       39.1 &       34.6 &       33.0 &       32.5 &       32.5 &       32.8 &       33.2 &       34.1 &       34.3 &       34.5 &       35.1 &       36.6 &       36.9 &       35.5 &       36.0 &       36.5 &       37.3 \\
         &     x100 &      795.0 &      262.1 &      140.9 &      104.3 &       90.4 &       81.2 &       77.2 &       75.1 &       69.7 &       70.9 &       66.5 &       65.6 &       64.9 &       64.0 &       58.9 &       58.1 &       60.2 &       58.4 &       57.5 &       55.8 \\
